---
title: 'NANOOS Profiles: SCA Shoreline Characterization'
author: "Regina Lionheart"
date: "2023-03-01"
output: pdf_document
editor_options: 
  markdown: 
    wrap: sentence
---

## Notes from analysis

The profiles that don't match up with WCEHA data are:

-   Profiles 14 and 15 at the North Jetty

    -   These ones are the most significant "at odds" data, so warrant a second look. NANOOS indicates significant accretion, while WCEHA indicates erosion.

-   Profiles 16 and 17 (Westport North and South)

    -   16 is Westport North, and WCEHA data says eroding, while NANOOS says accreting. It's true that Westport North is eroding overall, but WCEHA contains data from 2006 - 2019, and NANOOS contains data from 1997 - 2022. NANOOS encompasses the Army Corp nourishment more holistically. Both of the Westport profiles were "non significant" accretion, so I think that speaks to how much is moving around on that peninsula.

-   Profile 26, one of the Leadbetter Point profiles, which falls in a very small 'erosion' window on the WCEHA dataset, and is also the site of vegetation clearance for snowy plover habitat.

-   Profile 50, located at Seaview.
    NANOOS says erosion but WCEHA says accretion, however, the NANOOS data is non significant.

-   Profile 51, just north of the North Head lighthouse.
    This is the opposite of 50, where NANOOS says accretion but WCEHA says erosion, and also is a non-significant data point from the NANOOS profile.

*Profiles 42 - 47 are in Oregon, and so have been excluded from much of the analysis.*

*Profiles do not proceed sequentially.*

# Overview

NANOOS Shoreline profiles in x y z format (Easting Northing elevation).
Naming convention is prof_X\_ttYY.out where prof is short for profile, X is the profile number, tt is a season code (e.g. f = fall) and YY is the year, in the format of the last two digits (i.e, "98" is 1998", "00" is 2000, "08" is 2008, etc).

Quite a few profiles are null (empty) and will be removed and noted.
See the "Explore Profiles" script for more information.

Some of the profiles had non-conforming names: "BigE06", "beachface", etc.
These do not represent a large portion of the files and have not been included in the analysis.

Two duplicates were removed from original set: see accompanying script "duplicate_test.R".

### Metadata

The vertical datum is NAVD88.
The horizontal is WA State Plane South.
All of the units are in meters.

### Script setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      results = "hide")
knitr::opts_hooks$set(label = function(options) {
   if (is.null(options$fig.cap)) {
    stop("Every figure has to have a caption!")  
   }
   return(options)
})

knitr::opts_hooks$set(label = function(options) {
  default_label <- knitr::opts_knit$get("unnamed.chunk.label")
  has_default_label <- grepl(default_label, options$label)
  
  if (has_default_label) {
    stop("Name your chunks!")
  }
  return(options)
})

source("scripts/src/kableExtrafunction.R")
```

## Load required packages

```{r load packages, echo=TRUE, results='hide'}
source("scripts/src/load_packages.R")
```

## Explore profiles

```{r prep, echo=FALSE}
source("scripts/examine_profiles.R")
```

## View profiles and their locations

```{r prep, echo=FALSE, warning=FALSE}
knitr::kable(profiles.w.location, caption = "Profiles with matching geographic location") %>%
  kableExtra::kable_styling(fixed_thead = T)
```

## Individual profile, single year visualization

Select a single profile by entering "prof_N" below, with "N" being a profile number chosen from the above list.
Select a single year by entering the last two digits of the year (from 98 to 18).
The output from this code chunk will print the total number of files associated with the request, as well as any empty files within the request parameters, and plot the given profile.

This is an actual plot of the x y z coordinates to understand file format as well as the variability of a given profile, *not* a direct geographic representation of the data.
The graph can be rotated using the mouse.
Hover over the lines to see the seasonal data.

```{r prep, echo=FALSE}
profile.pattern <- "prof_1"
year.pattern <- c("99")

source("scripts/single_profile_3D.R")

suppressWarnings(single.profile.plot)
```

## Individual profile, all available years visualization

Import data for a single profile, automatically including all available years.
Create a visual for the complete profile with all years.
The orange horizontal plane is the MHHW mark.

```{r, echo=FALSE}
profile.pattern <- "prof_16"

source("scripts/complete_profile_plane.R")

suppressWarnings(print(complete.profile.plot))
```

It is also helpful to see the above plot spread over time, to provide a clearer picture of profile migration.

```{r, echo=FALSE}
profile.pattern <- "prof_23"

source("scripts/complete_profile_timeseries.R")

suppressWarnings(print(profile.timeseries.fig))

```

## "BasePoint" reference line and Euclidean distance as an accretion/erosion proxy

In order to quantify shoreline change, an arbitrary "BasePoint" reference line, parallel to the water, was drawn at a reasonable distance from the profiles.
The profiles (aka transects) are perpendicular to this BasePoint, allowing for a reference point against which to calculate spatial change.

In order to visualize this concept, the below chunk of code visualizes 1) a simplified selection of profiles (plotted in black) compared to the BasePoint (plotted in red) and 2) an individual profile (plotted in black) and its BasePoint (plotted in red).

```{r, echo=FALSE}
source("scripts/profiles_and_BasePoint.R")
print(all.basepoint.plot)
print(single.basepoint.plot)
```

Using Euclidean distance, migrations of the transects over time as compared to the BasePoint act as proxies for accretion and erosion.

|                               |           |
|-------------------------------|-----------|
| Increasing Euclidean distance | Accretion |
| Decreasing Euclidean distance | Erosion   |

## Characterizing profile elements to use for accretion and erosion proxies.

Find the midpoint of each profile to acquire a single data point to describe the overall migration of the profile.
See where that midpoint (and by extension the entire profile) is migrating with time.

```{r, echo=FALSE}
profile.pattern <- "prof" ## Adjust this to see specific profiles.
source("scripts/midpoint_euclidean_distance.R")
midpoint.euc.dist.plot
```

## Assign status to each profile according to linear regression equation details.

Once the midpoint euclidean distance has been plotted against time, a linear regression can be applied to understand how time is a predictor of erosion or accretion, and whether that profile movement is considered significant or not.

Below is a table with the relevant results of each profile's linear regression over time, including pvalues (less than 0.05 indicates a significant relationship), rsquared values (indicitive of how good the fit of the regression is) and the Pearson correlation (how closely are time and euclidean distance, or x and y, related).

```{r prep, echo=FALSE}
knitr::kable(results.table, caption = "Linear equation parameters") %>%
  kableExtra::kable_styling(fixed_thead = T)
```

## Comparison of results to Washington Coastal Erosion Hazard Assessment (WCEHA)

The WCEHA (<https://waecy.maps.arcgis.com/apps/View/index.html?appid=389d0a3ce642485db912d4a416a56e25>) is a second database that can be used as a comparison to the results from the NANOOS profiles.

```{r}
source("scripts/WCEHA_validation.R")
knitr::kable(all.data, caption = "Comparison with WCEHA data") %>%
  kableExtra::kable_styling(fixed_thead = T) %>%
  row_spec(which(all.data$conflict == "flag"), bold = T, color = "white",
           background = "tomato")

```

## Linear Regressions and Change Rates

The midpoint of each profile selects a single point as a proxy for erosion through time.
This obscures the variation inherent in the collection of profiles.

Therefore it is important to find a method of individual profile characterization that will apply to all 4500 profiles, and can act as a reliable proxy to determine rates of erosion or accretion.

The profiles have linear relationships due to the geographic layout: in general, a change in x (moving along the Easting plane) will result in a *roughly* equal change in y (the Northing plane).
Therefore, a linear regression can be applied to the x and y variables of each profile to obtain a line proxy rather than a single point.

Look at a single profile and year to see the applied linear regression.
On the right, the linear quartiles are added for more detailed data description of accretion and erosion.
These quartiles are labeled "minimum", "quartile1", "median", "quartile3", and "maximum" in accordance with standard boxplot parameters, although they are calculated using *only distance* and not boxplot delineations.

```{r, echo=FALSE}
profile.pattern <- "prof_7"
year.pattern <- "00"

source("scripts/single_linear_quartiles.R")
single.linear.plot + single.quartiles.plot
```

Characterizing the profiles with more points allows for a more detailed look at erosion and accretion rates.

Each quartile's movement is tracked over time, same as the midpoint proxy used above.
Then the means of the movement was taken and used to determine a rate of change for each profile.

The rates can be viewed by park or by profile.

```{r, echo=FALSE}
profile.pattern <- "prof"
source("scripts/rates_of_change.R")
region.ROC.plot
profile.ROC.plot
```

## Using Mean High High Water (MHHW) as a proxy for shoreline change

As the profiles (and the shoreline) change over time, the coordinate at which a profile crosses the mean high high water mark shifts, acting as a reliable indicator of accretion and erosion.

Visualize where the profiles cross the MHHW mark.

```{r, echo=FALSE}
source("scripts/MHHW_to_BasePoint.R")

complete.profile.plot
MHHW.to.BasePoint
```

How does the euclidean distance from the MHHW to the BasePoint move with time?
What are the rates of change when the proxy is the MHHW mark?

```{r, echo=FALSE}
profile.pattern <- "prof"
source("scripts/MHHW_RatesOfChange.R")
MHHW.ROC.plot
```

Clustering using the midpoint euclidean distance data.

This analysis applies unsupervised clustering, as we want are inputting unlabeled data without training.
Relevant types of unsupervised clustering are k-means clustering and hierarchical agglomerative clustering (HCA).

Pure kmeans requires some advance knowledge of the number of clusters k, while HCA seeks to build a hierarchy of clusters without a predetermined k.

Because we are trying to determine the sub reaches, we will use HCA for this analysis.
The number of classes is not specified in advance; hierarchical clustering will determine that.

Our variables are distance to baseline (euclidean midpoint movement over time, aka erosion and accretion), and year.

HCA Approach 1.
Create a distance matrix that computes and returns the distance between rows.
2.
Start with every point in its own cluster and compare distances.
3.
Merge the two points that are closest to each other based on "distance" (distance method is variable), from the calculated distance matrix in step 1.
This removes one cluster.
This also corresponds to the first height of the luster dendogram.
- Measure cluster distance by single linkage: compute minimum distance before merging.
- "Single" is the default distance measure but may not be the most effective.
- "Single" also falls under the umbrella of nearest neighbor method, which builds branches off of smallest distances.
4.
Recalculate all distances with the new number of clusters and continue to combine.
5.
Continue this until all clusters have been merged.
6.
Create a dendogram using the history of the point combination, which shows the sequence of combinations of the clusters.
The distances of merge between clusters, called heights, are illustrated on the y-axis.

```{r, echo=FALSE}
source("scripts/cluster_analysis.R")

```

## Final Output for Attribute Table

This is the attribute table csv for the final product of the interactive GIS map.
The table includes:

1.  Profile

2.  Geographic Location

3.  Overall movement of each (euclidean distance)

4.  Designation of accretion/erosion and accompanying significance

5.  Rates of change of each

6.  Subreach delineated by clustering

Things that I would like to include if I have a few more hours:

1.  Inflection points

2.  Quality of profile data

```{r, echo=FALSE}
source("scripts/final_output.R")

```
