---
title: 'NANOOS Profiles: SCA Shoreline Characterization'
author: "Regina Lionheart"
date: "2023-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Shoreline profiles in x y z format (easting northing elevation).
Named prof_X_ttYY.out where X is the profile number, tt is a season code (e.g. f = fall) and YY is the year.
The vertical datum is NAVD88.
The horizontal is WA State Plane South. All of the units are in meters.

Quite a few profiles are null and will be removed and noted.
We have two duplicates, see accompanying script "duplicate_test.R"
Check on non-conforming files: beachface? dunes?


Load packages
```{r prep, echo=FALSE}
source("scripts/src/load_packages.R")
```

## View individual profile and year.
Select a profile of interest by entering "prof_N" below. 

Select a year of interest by entering the last two digits of the year (from 98 to 18).
```{r prep, echo=FALSE}
profile.pattern <- "prof_6"
year.pattern <- c("00")

source("scripts/src/import_profiles.R")
```

Visualize the single profile/year.
```{r prep, echo=FALSE}
source("scripts/single_profile_3D.R")
profileplot
```


View complete profile including all years.
