---
title: "21-07771-003: Shoreline Change Analysis"
author: "Regina Lionheart"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
palette =c("#04A1FF",'#FFC42E','#3ECDA3','#1D455C','#5C4E2C','#395C51')
```

## R Markdown

Tags
quartiles


> Herrera Environmental Consultants


## Table of Contents

-   [Project Description](#-Project-Description)
    -   [Data Discrepancies](#-Data-Discrepancies)
    -   [Metadata](#-Metadata)
-   [Location in Herrera Database](#-Location-in-Herrera-Database)  
-   [Requirements](#-Requirements)
-   [Usage](#-Usage)
    -   [Layout of Directory and Data](#-Layout-of-Directory-and-Data)
-   [Visualizations](#-Visualization)


------------------------------------------------------------------------

### Description {#description}

**Important:** *Link to VantagePoint somehow here*

This analysis aims to characterize approximately 25 years of shoreline profile transect data, provided by NANOOS (Northwest Association of Networked Ocean Observing Systems).
Using this data combined with area expertise, the relevant coastal areas have been delineated into reaches.

The raw data consists of approximately 4500 individual profiles across 54 coastal sites in Washington and Oregon.
The profiles are in x y z format (Easting Northing elevation).The naming convention is prof_X_ssYY.out where prof is short for profile, X is the profile number, ss is a season code (e.g. f = fall) and YY is the year, in the format of the last two digits (e.g. "98" is 1998", "00" is 2000, "08" is 2008, etc).

#### Data discrepancies {#data-discrepancies}

Quite a few profiles are null (empty) and have been removed and noted.
See the "Explore Profiles" section for more information.

A small portion of the profiles had non-conforming filenames: "BigE06", "beachface", etc.
These files do not represent a large percentage of the files and have not been included in the analysis.

Profiles 42 - 47 are in Oregon and have been excluded from most of the analysis.

*Geographically, profiles do not proceed sequentially.*

#### Metadata {#metadata}

The vertical datum is NAVD88.
The horizontal is WA State Plane South.
All of the units are in meters.


### 📦 Requirements {#requirements}

Below is a list of packages and external sofwares that this project utilizes. Please ensure you have the package(s) installed and have access to the tools listed below.

```{r packages, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
packages.table <- "
|Name          | Description           
|:-------------|:-------------|
|[`R`](https://www.r-project.org/) | Programming language used for this project. |
|[`NANOOS Profile Data`](https://nvs.nanoos.org/BeachMapping) | The raw data for analysis. If you do not have access to this data, please contact the owners of this repository.|
|[`BasePoint Coordinates`] | Easting Northing coordinates for the arbitrary basepoint line.|
"
cat(packages.table)
```

If you do not have access to the data, please contact the emails listed at the bottom of the repository.

### Usage {#usage}

The first few chunks of the Control.Rmd script will create empty folders for different types of data (raw, processed, extra, intermediate). Some of this data will be produced by the script (processed, intermediate) and some needs to be moved into those folders before proceeding. Because this is a template script, the data has been modified for easy input, which will NEVER be the case in a real run.

All the necessary example data can be downloaded here: <https://drive.google.com/drive/folders/1E-B4dyDTkwOfupHJK0uL8BgKj586OU31>

If you do not have access to this drive, please contact the owners of this repository.

> #### User-Defined QC Parameters {#user-defined-QC-paramters}

These parameters will change depending on your data and level of stringency. Below are the "default" values.

**area.min**: The minimum area of an integrated peak to be considered a real peak.
HILIC - 1000, Reverse Phase - 5000.

**RT.flex**: Flexibility of the deviation of the sample retention time from the appropriate standard. HILIC +/- 0.4 min, Reverse Phase +/- 0.2 min.

**blk.thresh**: Cutoff ratio value of sample area to a blank reference sample. HILIC +/-0.3, Reverse Phase +/- 0.2.

**SN.min**: Cutoff ratio value for Signal to Noise. HILIC +/- 4, Reverse Phase +/- 4.

#### Additional QC parameters for Skyline 

**height.min**: Minimum height for a peak. 1000, HILIC and Reverse Phase.

**height.max**: Maximum height for a peak. 1.0e8, HILIC and Reverse Phase.

#### Additional QC parameter for TQS 

**IR.flex**: The TQS produces quantitative and secondary trace values. Find the Ion Ratio by dividing the area of the quantitative trace by the area of the secondary trace for standards, and use these ratios to create a reference table. Deviations from this reference table that are larger than the user-defined values will be flagged. 0.3, HILIC and Reverse Phase.

## Visualization {#visualization}

![Click on me to see a visual layout of the Targeted Pipeline!](visual/Targeted_Pipeline_Visualization.pdf)

------------------------------------------------------------------------


## 🔧 Pull Requests

Pull requests are welcome. For major changes, please open an issue first.

## 💬 Contributors + Contact Information

-   [Regina Lionheart](https://github.com/R-Lionheart)

------------------------------------------------------------------------

![dumbbell_plot](https://img.shields.io/badge/plot-dumbbell-%231D455C) 
![HCA_clustering](https://img.shields.io/badge/stats-HCA%20clustering-%233ECDA3)
![linear_model](https://img.shields.io/badge/stats-linear%20model-%231D455C)
![boxplot](https://img.shields.io/badge/plot-boxplot-%231D455C)
![3D_regression](https://img.shields.io/badge/plot-3D%20regression-%231D455C)
![quartiles](https://img.shields.io/badge/stats-quartiles-%233ECDA3)
